\documentclass{article}

\usepackage{fullpage}

\begin{document}

\section{Problem definition}

Graphs. $G=(V,E)$

We are looking at stencil graphs. 2D stencil graphs with different
neighborhood.  The vertices are numbered with a coordinate $(i,j)$. So
a $N \times M$ stencil has vertices for $(0,0), (0,1), \dots (0, N-1),
(1,0), (1,1), \dots (M-1, N-1)$.  We are considering three different
neighborhood.

\paragraph{5-point stencil} where vertex $(i,j)$ is connected to $(i-1,j)$,
$(i+1,j)$, $(i,j-1)$, $(i,j+1)$

\paragraph{9-point box stencil} where vertex $(i,j)$ is connected to
$(i-1,j-1)$, $(i-1,j)$, $(i-1,j+1)$,
$(i,j-1)$, $(i,j+1)$,
$(i+1,j-1)$, $(i+1,j)$, $(i+1,j+1)$.

\paragraph{9-point star stencil} where vertex $(i,j)$ is connected to
$(i-2,j)$, $(i-1,j)$, $(i,j-1)$, $(i,j-2)$, $(i,j+1)$, $(i,j+2)$,
$(i+1,j)$, $(i+2,j)$.

All of these neighborhoods are assuming the neighbors exist (are
within the $N \times M$ stencil).


The problem we are looking at is \textbf{star coloring}. A star coloring of a
graph is a function on the vertices $C: V \rightarrow N$ such that,
$\forall (u,v) \in E, C(u) \neq C(v)$ and such that every chain of 4
distinct vertices in the graph has at least 3 colors.

An equivalent (?) definition of a star coloring is that the graph
induced by the ertices of any two colors is a forest of star graphs.

We want to find the star coloring that uses the fewer number of colors.

\section{ILP}

Wrote the ILP to solve the problem optimally. Here is what we know
from the ILP.

On 5-pt stencil. The optimal star coloring of a $3 \times 3$ stencil
has 4 colors, while $4 \times 4$ needs 5 colors. YOu can easily find a
5 color solution for 5-pt stencil. For instance, you can color a 5pt stencil
centered on a vertex with 5 different colors and then tile the plane
with that pattern to make a valid 5 color star coloring.

On 9-pt box stencil. The $3 \times 3$ stencil needs 5 colors, the $4
\times 4$ stencil needs 7 colors, and from sizes of $5 \times 5$,
there are 8-color solutions up to stencils of $48 \times 48$.

On 9-pt star stencil. The $3 \times 3$ stencil and $4 \times 4$
stencil need 6 colors. The $5 \times 5$ stencil and $6 \times 6$
stencil need 7 colors. The $7 \times 7$ stencil, $8 \times 8$ stencil,
and $9 \times 9$ stencil need 8 colors.

\section{All 9pt-box stencils are 8 colorable}


\end{document}
